---
title: "Practical Machine Learning Assignment"
author: "Peter Whitney"
date: "January 24, 2016"
output: html_document
---

# Synopsis
For this assignment we will perform an analysis with predictions on the 
"Qualitative Activity Recognition of Weight Lifting Exercises"
dataset referenced via the following URL: 
"http://groupware.les.inf.puc-rio.br/har#dataset"

# Data Preparation
The referenced dataset includes many rows that both add new features and
calculated variables. The added variables and features are not part of the
initial data collection set and adding these variables and features makes 
the dataset a non-tidy dataset. As a result, all of the features designated 
by the new_window="yes" designation will be removed. Additionally all of the 
dervied colums will also be purged. The resulting dataset is one that is tidy, 
and fully populated without any need to extrapolate or fill in missing data points.

All of the same operations are also performed on the test dataset.
```{r}
setwd("/Users/pwhitney/Documents/Coursera/PracticalMachineLearning")
train_initial <- read.csv("pml-training.csv")
train_purge_calculated <- train_initial[train_initial$new_window=="no",]
train_cols <- c("user_name",
		"accel_arm_x","accel_arm_y","accel_arm_z",
		"gyros_arm_x","gyros_arm_y","gyros_arm_z",
		"magnet_arm_x","magnet_arm_y","magnet_arm_z",
		"pitch_arm","roll_arm","yaw_arm","total_accel_arm",
		"accel_belt_x","accel_belt_y","accel_belt_z",
		"gyros_belt_x","gyros_belt_y","gyros_belt_z",
		"magnet_belt_x","magnet_belt_y","magnet_belt_z",
		"pitch_belt","roll_belt","yaw_belt","total_accel_belt",
		"accel_dumbbell_x","accel_dumbbell_y","accel_dumbbell_z",
		"gyros_dumbbell_x","gyros_dumbbell_y","gyros_dumbbell_z",
		"magnet_dumbbell_x","magnet_dumbbell_y","magnet_dumbbell_z",
		"pitch_dumbbell","roll_dumbbell","yaw_dumbbell","total_accel_dumbbell",
		"accel_forearm_x","accel_forearm_y","accel_forearm_z",
		"gyros_forearm_x","gyros_forearm_y","gyros_forearm_z",
		"magnet_forearm_x","magnet_forearm_y","magnet_forearm_z",
		"pitch_forearm","roll_forearm","yaw_forearm","total_accel_forearm", "classe")
train <- subset(train_purge_calculated, select=train_cols)

test_initial <- read.csv("pml-testing.csv")
test_purge_calculated <- test_initial[test_initial$new_window=="no",]
test_cols <- train_cols[1:53]
test <- subset(test_purge_calculated, select=test_cols)
```

# Model Generation
I chose to use Random Forests with 10 Fold Cross Validation for this
assignment. Random Forests was chosen for its known ability to produce
good classification result across a wide variety of datasets and it is
the same choice used by the study team that performed the initial study.

I did not apply any ensembling methods as the results I achieved with just
Random Forests came so close to the initial study results that pursuing
further improvments in accuracy would have added more complexity and execution
time and produced very little improvement if any at all. You can think of this
as being similar to the NetFlix contest where award winning results were achieved
but the model was never implemented due to execution time and complexity considerations.
```{r}
require(caret)
require(randomForest)

set.seed(12345)
rf_model<-train(classe ~ .,data=train,method="rf",
                trControl=trainControl(method="cv",number=10),
                prox=TRUE,allowParallel=TRUE)
```

# Random Forests Results
The results indicate that one model was able to accomplish accuracy performance
of 99.5% Which is an exceptional accuracy result by many standards and comes very
close to the results achived by the study group.
```{r}
print(rf_model)
```

# Random Forests Results and Estimated Error Rate
The estimated error rate is 0.44%.
```{r}
print(rf_model$finalModel)
```

# Create Test Set Predictions Using the Random Forest Model
Note that the predictions created below cannot be that bad as I was able to make
ALL quiz predictions correctly (100%) using the process shown above.
```{r}
rf_predict <- predict(rf_model, test)
for(i in 1:20){
    line = paste0("prediction_",i," : ",rf_predict[i])
    print(line)
}
```
